---
title: "Replication of Petersen (2009) Simulation"
output: rmarkdown::github_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(parallel)
library(foreach)
library(doParallel)
library(lmtest)
library(sandwich)
library(plm)
library(knitr)
library(tidyverse)
set.seed(5, kind = "L'Ecuyer-CMRG")
```

[Petersen (2009)](https://www.kellogg.northwestern.edu/faculty/petersen/htm/papers/standarderror.html) is an influential paper guiding researchers picking correct standard errors when using linear panel models (for the case without Omitted Variable Bias (OVB)). A brief summary of the paper's results can be found in [my study note of linear panel model (Section 7.1)](https://github.com/maxleungtszchun/My-Econometrics-Notes/blob/main/notes%20on%20linear%20panel%20model.pdf).

Here is my R script replicating the paper's key results. It can be executed in both Windows and MacOS. However, it will be run much faster in MacOS by using multiple CPUs due to the use of `doParallel` package. To reduce running time, only 500 simulations instead of 5000 simulations executed in the paper are performed.

```{r results = "asis"}
gen_long_data <- function (N, T, var_gamma_i, var_mu_i, fe = "individual") {
  d <- tibble(.rows = N) %>% mutate(i = seq_len(N))
  d["gamma_i"] <- rnorm(N, mean = 0, sd = sqrt(var_gamma_i))
  d["mu_i"] <- rnorm(N, mean = 0, sd = sqrt(var_mu_i))
  
  gen_wide_data_i_eff <- function (df) {
    var_eta_it <- 4 - var_gamma_i
    var_upsilon_it <- 1 - var_mu_i
    for (t in seq_len(T)) {
      df[str_c("eta_i", t)] <- rnorm(N, mean = 0, sd = sqrt(var_eta_it))
      df[str_c("varepsilon_i", t)] <- df[["gamma_i"]] + df[[str_c("eta_i", t)]]
      df[str_c("upsilon_i", t)] <- rnorm(N, mean = 0, sd = sqrt(var_upsilon_it))
      df[str_c("x_i", t)] <- df[["mu_i"]] + df[[str_c("upsilon_i", t)]]
      df[str_c("y_i", t)] <- df[[str_c("x_i", t)]] * 1 + df[[str_c("varepsilon_i", t)]]
    }
    return(df)
  }
  
  gen_wide_data_t_eff <- function (df) {
    var_delta_t <- var_gamma_i
    var_zeta_t <- var_mu_i
    
    var_eta_it <- 4 - var_delta_t
    var_upsilon_it <- 1 - var_zeta_t
    for (t in seq_len(T)) {
      df[str_c("eta_i", t)] <- rnorm(N, mean = 0, sd = sqrt(var_eta_it))
      df[str_c("varepsilon_i", t)] <- rnorm(1, mean = 0, sd = sqrt(var_delta_t)) +
                                        df[[str_c("eta_i", t)]]
      df[str_c("upsilon_i", t)] <- rnorm(N, mean = 0, sd = sqrt(var_upsilon_it))
      df[str_c("x_i", t)] <- rnorm(1, mean = 0, sd = sqrt(var_zeta_t)) +
                              df[[str_c("upsilon_i", t)]]
      df[str_c("y_i", t)] <- df[[str_c("x_i", t)]] * 1 + df[[str_c("varepsilon_i", t)]]
    }
    return(df)
  }
 
  gen_wide_data_it_eff <- function (df) {
    var_delta_t <- var_gamma_i
    var_zeta_t <- var_mu_i
    
    var_eta_it <- var_gamma_i
    var_upsilon_it <- var_mu_i
    for (t in seq_len(T)) {
      df[str_c("eta_i", t)] <- rnorm(N, mean = 0, sd = sqrt(var_eta_it))
      df[str_c("varepsilon_i", t)] <- df[["gamma_i"]] +
                                      rnorm(1, mean = 0, sd = sqrt(var_delta_t)) +
                                      df[[str_c("eta_i", t)]]
      df[str_c("upsilon_i", t)] <- rnorm(N, mean = 0, sd = sqrt(var_upsilon_it))
      df[str_c("x_i", t)] <- df[["mu_i"]] +
                              rnorm(1, mean = 0, sd = sqrt(var_zeta_t)) +
                              df[[str_c("upsilon_i", t)]]
      df[str_c("y_i", t)] <- df[[str_c("x_i", t)]] * 1 + df[[str_c("varepsilon_i", t)]]
    }
    return(df)
  }
  
  wide_data <- 
    switch(fe,
           individual = gen_wide_data_i_eff(d),
           time = gen_wide_data_t_eff(d),
           both = gen_wide_data_it_eff(d),
           stop("only individual or time or both fe"))
  long_data <-
    wide_data %>%
    pivot_longer(cols = starts_with(c("y_i","x_i", "varepsilon_i",
                                      "eta_i", "upsilon_i")),
                 names_to = c(".value", "t"),
                 names_sep = "_i")
  return(long_data)
}

simulation <- function (N, T, var_gamma_i, var_mu_i, sim_num, fe = "individual") {
  getResults <- function() {
    long_data <- 
      gen_long_data(N = N,
                    T = T,
                    var_gamma_i = var_gamma_i,
                    var_mu_i = var_mu_i,
                    fe = fe) %>%
      mutate(t = as.numeric(t))
    long_data_plm <- pdata.frame(long_data, index = c("i", "t"))
    
    pooled_ols <- plm(y ~ x + 0, model = "pooling", data = long_data_plm)
    ols <- lm(y ~ x + 0, data = long_data) # for calculating White's S.E. later by using sandwich::vcovHC()
    pooled_ols_coef_default_se <- coeftest(pooled_ols)
    clustered_i_var <- plm::vcovHC(pooled_ols, type = "HC1", cluster = "group")
    pooled_ols_coef_clustered_se_i <- coeftest(pooled_ols, vcov = clustered_i_var)
    clustered_t_var <- plm::vcovHC(pooled_ols, type = "HC1", cluster = "time")
    pooled_ols_coef_clustered_se_t <- coeftest(pooled_ols, vcov = clustered_t_var)
    # plm::vcovHC() =/= sandwich::vcovHC()
    clustered_both_var <- clustered_i_var + clustered_t_var - sandwich::vcovHC(ols, type = "HC1")
    pooled_ols_coef_clustered_se_both <- coeftest(pooled_ols, vcov = clustered_both_var)
    
    pooled_ols_coef_clustered_se <-
      switch(fe,
             individual = pooled_ols_coef_clustered_se_i,
             time = pooled_ols_coef_clustered_se_t,
             both = pooled_ols_coef_clustered_se_both,
             stop("only individual or time or both fe"))
    
    cs_betas <- list()
    for (j in seq_len(T)) {
      cs_betas[[j]] <-
        long_data %>%
        filter(t == j) %>%
        lm(y ~ x + 0, data = .) %>%
        summary(.) %>%
        .$coefficients %>%
        .[1,"Estimate"]
    }
    cs_betas <- as.numeric(cs_betas)
    fm_beta <- mean(cs_betas, na.rm = TRUE)
    fm_se <- sqrt(var(cs_betas, na.rm = TRUE) / T)
    
    return(list("default" = pooled_ols_coef_default_se,
                "clustered" = pooled_ols_coef_clustered_se,
                "fm_beta" = fm_beta,
                "fm_se" = fm_se,
                "fm_t" = (fm_beta - 1) / fm_se))
  }
  
  if (Sys.info()["sysname"] == "Windows") {
    results <- list()
    for (i in seq_len(sim_num)) {
      results[[i]] <- getResults()
    }
  } else {
    # using all CPUs in non-Windows OS e.g. MacOS
    registerDoParallel(detectCores())
    results <-
      foreach (i = seq_len(sim_num)) %dopar% {
        getResults()
      }
    stopImplicitCluster()
  }
  
  return(results)
}

get_sim_table <- function (N, T, var_gamma_i, var_mu_i, sim_num, fe = "individual") {
  sim_data <- simulation(N = N,
                         T = T,
                         var_gamma_i = var_gamma_i,
                         var_mu_i = var_mu_i,
                         sim_num = sim_num,
                         fe = fe)
  
  t_default <- map_dbl(sim_data, ~ .x$default[1,"t value"] - 1 / .x$default[1,"Std. Error"])
  t_cluster <- map_dbl(sim_data, ~ .x$clustered[1,"t value"] - 1 / .x$clustered[1,"Std. Error"])
  t_fm <- map_dbl(sim_data, ~ .x$fm_t)
  
  per_sig_t_ols <- mean(as.numeric(abs(t_default) > 2.58), na.rm = TRUE)
  per_sig_t_c <- mean(as.numeric(abs(t_cluster) > 2.58), na.rm = TRUE)
  per_sig_t_fm <- mean(as.numeric(abs(t_fm) > 2.58), na.rm = TRUE)

  if (var_gamma_i == 2 & var_mu_i == 0.5) {
    plotHist <- function (d, title) {
      # dev.new()
      hist(d, prob = TRUE, xlim = c(-5, 5), ylim = c(0, 1), main = title, col = "orange", "border" = "white")
      # xfit <- seq(min(d, na.rm = TRUE), max(d, na.rm = TRUE), length = 200)
      # yfit <- dnorm(xfit,
      #               mean = mean(d, na.rm = TRUE),
      #               sd = sd(d, na.rm = TRUE))
      # lines(xfit, yfit, col = "red")
    }
    plotHist(t_default, "Density of simulated t with default standard error")
    plotHist(t_cluster, "Density of simulated t with clustered standard error")
    plotHist(t_fm, "Density of simulated t with Fama-Macbeth standard error")
  }

  simulated_pooled_ols_betas <- map_dbl(sim_data, ~ .x$default[ ,"Estimate"])
  return(list("avg_beta_ols" = mean(simulated_pooled_ols_betas, na.rm = TRUE),
              "std_beta_ols" = sd(simulated_pooled_ols_betas, na.rm = TRUE),
              "avg_se_ols" = mean(map_dbl(sim_data, ~ .x$default[ ,"Std. Error"]), na.rm = TRUE),
              "per_sig_t_ols" = per_sig_t_ols,
              "avg_se_c" = mean(map_dbl(sim_data, ~ .x$cluster[ ,"Std. Error"]), na.rm = TRUE),
              "per_sig_t_c" = per_sig_t_c,
              "avg_beta_fm" = mean(map_dbl(sim_data, ~ .x$fm_beta), na.rm = TRUE),
              "std_beta_fm" = sd(map_dbl(sim_data, ~ .x$fm_beta), na.rm = TRUE),
              "avg_se_fm" = mean(map_dbl(sim_data, ~ .x$fm_se), na.rm = TRUE),
              "per_sig_t_fm" = per_sig_t_fm))
}

# Var(varepsilon_it) = 4
# Var(x_it) = 1
sim_num <- 500
N <- 500
T <- 10
var_gamma_var_mu_list <- 
  list(list(0, 0),
       list(0, 0.25),
       list(0, 0.5),
       list(0, 0.75),
       list(1, 0),
       list(1, 0.25),
       list(1, 0.5),
       list(1, 0.75),
       list(2, 0),
       list(2, 0.25),
       list(2, 0.5),
       list(2, 0.75),
       list(3, 0),
       list(3, 0.25),
       list(3, 0.5),
       list(3, 0.75))

start_time <- Sys.time()
sim_table_i <-
  var_gamma_var_mu_list %>%
  map( ~ get_sim_table(N = N, T = T, var_gamma_i = .x[[1]],
                       var_mu_i = .x[[2]], sim_num = sim_num)) %>%
  do.call(bind_rows, .) %>%
  t()
colnames(sim_table_i) <- var_gamma_var_mu_list
kable(sim_table_i, caption = "Only Individual Fixed Effect")

# for time fixed effect,
# var_gamma_i in get_sim_table() is actually var_delta_t, and
# var_mu_i in get_sim_table() is actually var_zeta_t
sim_table_t <-
  var_gamma_var_mu_list %>%
  map( ~ get_sim_table(N = N, T = T, var_gamma_i = .x[[1]],
                       var_mu_i = .x[[2]], sim_num = sim_num, fe = "time")) %>%
  do.call(bind_rows, .) %>%
  t()
colnames(sim_table_t) <- var_gamma_var_mu_list
kable(sim_table_t, caption = "Only Time Fixed Effect")

sim_table_both <-
  list(list(4/3, 1/3)) %>%
  map( ~ get_sim_table(N = 100, T = 100, var_gamma_i = .x[[1]],
                       var_mu_i = .x[[2]], sim_num = sim_num, fe = "both")) %>%
  do.call(bind_rows, .) %>%
  t()
kable(sim_table_both, caption = "Both Individual and Time Fixed Effect")
Sys.time() - start_time
```